---
title: "HW 04"
author: "Lucas Smith"
format:
  html:
    embed-resources: true
toc: true
---

```{r}

packages <- c("dplyr", "ggplot2", "tigris", "readxl", "ggrepel", "patchwork", "tidyverse", "grid", "gridExtra", "ggplot2", "reshape", "ggalt", "likert", "ggstats", "ggimage", "jpeg" ,"png")

for (pkg in packages) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
}

```


```{r load-packages, message = FALSE}
library(tidyverse)
library(tigris)
library(sf)
library(readxl)
library(patchwork)
library(ggrepel)
# add any other packages you use here
```


## 1 - A second chance

I'm taking from homework 3 - DuBois. To compare, here's the original plot I made. I chose this plot because it was not only challenging, but there were some things that I wanted to improve on. Getting the colors and everything lined up was a challenging aspect, but I think that making the final touches on it was something that I just didn't have time for. 

```{r}
#| label: dubois
library(jpeg)
parchment <- readJPEG("images/parchment.jpg")
bg <- rasterGrob(parchment, width = unit(1,"npc"), height = unit(1,"npc"))

dubois <- read_csv("data/income.csv")
db <- pivot_longer(dubois, 
                   c("Rent", "Food", "Clothes", "Tax", "Other"), 
                   names_to="income_type", values_to = "percent")
order <- rev(c("$100-200", 
               "$200-300", 
               "$300-400", 
               "$400-500", 
               "$500-750", 
               "$750-1000", 
               "$1000 AND OVER"
               )
             )
color_vals = c("Rent" = "#141512", "Food" = "#805D7D", 
               "Clothes" = "#DA907F", "Tax"="#958c82", 
               "Other"="#BDB39C"
               )
x_ordering <- c("Rent", "Food", "Clothes", "Tax", "Other")

plt <- ggplot(db, 
              aes(x=percent, 
                  y=Class, 
                  fill=factor(income_type, 
                              levels=rev(x_ordering)
                              ),
                  )) +    
  geom_col(position="stack") + 
  geom_text(aes(label=percent), position = position_stack(vjust = 0.5),
            size=3)

plt <- plt + 
        guides(fill=FALSE) + 
        theme(axis.text.x=element_blank(),
             axis.title.x=element_blank(),
             axis.ticks.x=element_blank(), 
             axis.ticks.y = element_blank(),
             legend.position = NULL, 
             plot.margin=margin(0,0,0,0),
             plot.background = element_blank(),
             axis.text.y = element_blank(),
             panel.background=element_blank(),
             panel.grid.minor=element_blank(),
             panel.grid.major=element_blank()) + 
        labs(x="", y="") +
        scale_y_discrete(limits=order) +
        scale_fill_manual(values=color_vals) 

tabledata <- dubois |> select(Class, Average_Income)
tab <- tableGrob(tabledata, rows=NULL, theme=ttheme_default(base_size = 7, 
                                                           panel.background=element_blank()))
tab$heights <- unit(rep(1/nrow(tab), nrow(tab) ), "npc")

data <- grid.arrange(
  arrangeGrob(tab, nullGrob(), heights=c(200, 5)),
  arrangeGrob(nullGrob(), plt, heights=c(4, 30)),
  widths=c(3, 8)
) 

grid.raster(parchment, width=unit(1, "npc"), height=unit(1, "npc"))
grid.draw(data)

```
To adjust, I:
- Added the percent sign
- Removed the 0's
- Fixed the black on black text!

```{r}
#| label: dubois revised
library(jpeg)
parchment <- readJPEG("images/parchment.jpg")
bg <- rasterGrob(parchment, width = unit(1,"npc"), height = unit(1,"npc"))

dubois <- read_csv("data/income.csv")
db <- pivot_longer(dubois, 
                   c("Rent", "Food", "Clothes", "Tax", "Other"), 
                   names_to="income_type", values_to = "percent") |> filter(percent > 1)


order <- rev(c("$100-200", 
               "$200-300", 
               "$300-400", 
               "$400-500", 
               "$500-750", 
               "$750-1000", 
               "$1000 AND OVER"
               )
             )
color_vals = c("Rent" = "#141512", "Food" = "#805D7D", 
               "Clothes" = "#DA907F", "Tax"="#958c82", 
               "Other"="#BDB39C"
               )
x_ordering <- c("Rent", "Food", "Clothes", "Tax", "Other")

plt <- ggplot(db, 
              aes(x=percent, 
                  y=Class, 
                  fill=factor(income_type, 
                              levels=rev(x_ordering)
                              ),
                  )) +    
  geom_col(position="stack") + 
   geom_text(
    aes(
      label = paste0(round(percent, 0), "%"),
      color = ifelse(income_type == "Rent", "white", "black")
    ),
    position = position_stack(vjust = 0.5),
    size = 3
  ) + scale_color_identity()

plt <- plt + 
        guides(fill=FALSE) + 
        theme(axis.text.x=element_blank(),
             axis.title.x=element_blank(),
             axis.ticks.x=element_blank(), 
             axis.ticks.y = element_blank(),
             legend.position = NULL, 
             plot.margin=margin(0,0,0,0),
             plot.background = element_blank(),
             axis.text.y = element_blank(),
             panel.background=element_blank(),
             panel.grid.minor=element_blank(),
             panel.grid.major=element_blank()) + 
        labs(x="", y="") +
        scale_y_discrete(limits=order) +
        scale_fill_manual(values=color_vals) 

tabledata <- dubois |> select(Class, Average_Income)
tab <- tableGrob(tabledata, rows=NULL, theme=ttheme_default(base_size = 7, 
                                                           panel.background=element_blank()))
tab$heights <- unit(rep(1/nrow(tab), nrow(tab) ), "npc")

data <- grid.arrange(
  arrangeGrob(tab, nullGrob(), heights=c(200, 5)),
  arrangeGrob(nullGrob(), plt, heights=c(4, 30)),
  widths=c(3, 8)
) 

grid.raster(parchment, width=unit(1, "npc"), height=unit(1, "npc"))
grid.draw(data)

```


## 2.  Arizona state of counties

```{r}
#| label: label-me-2

azcounties <- counties("Arizona")

```

```{r}
#| label: graph for q2
ggplot(azcounties) + 
  geom_sf() + 
  geom_label(aes(label=NAME, geometry=geometry), stat="sf_coordinates") +
  theme_minimal() + 
  labs(
    x="Longitutde", y="Latitude", title="Counties in Arizona State",
    caption="Source: Shapefile obtained using {tigris} R package, v2.0.1"
  ) + 
   theme(
    plot.caption = element_text(hjust = 0, size = 8)
  )

```

## 3.  Arizona state of population change

```{r}
#| label: graph for q3

csvdata <- read_excel("./data/co-est2023-pop-04.xlsx", skip=5, n_max=15, col_names=c(
  "county", "totalPopulation", "2020", "2021", "2022", "2023"
))
csvdata <- csvdata |> 
  mutate(total_pop_change_20_23 = `2023`-`2020`,
         county_name =sub(" County,.*", "", county),
         county=sub(".", "", county_name),
         total_pop_change_20_23 = case_when (
           total_pop_change_20_23 > 0 ~ total_pop_change_20_23,
           total_pop_change_20_23 <= 0 ~ 0
         )
)



joined <- inner_join(azcounties, csvdata, join_by(NAME==county))

maxi <- max(joined$total_pop_change_20_23)
mini <- min(joined$total_pop_change_20_23)
ggplot(
  joined, aes(geometry=geometry, fill=total_pop_change_20_23)
) + geom_sf( aes(label=NAME, geometry=geometry), color="white") + 
  theme_minimal() + 
  labs(
    x="Longitutde", y="Latitude", title="Resident Population Change for Counties in AZ",
    subtitle="July 01, 2020 to July 01, 2023",
    caption="Source: Shapefile obtained using {tigris R package, v2.0.1\npopulation change data from the US Census Bureau",
    fill = "Population Change"
  )  + 
  scale_fill_gradient2(midpoint=(maxi - mini) / 2, low="dodgerblue3", high="firebrick") 

```

## 4.  Arizona state of Indiginous Tribal Regions

```{r}
#| label: reading data for q4
a1 <- read_sf("./data/American_Indian_Reservations_in_Arizona.shp")

a2 <- a1 |> filter(
  NAME %in% c("Hopi Tribe", "Navajo Nation", "White Mountain Apache Tribe", "San Carlos Apache Tribe", "Tohono O’odham Nation")
)


```



```{r}
#| label: visualizing q4
ggplot(azcounties) + geom_sf() + geom_sf(data=a1, aes(label=NAME), linewidth=1, fill="lightgrey", color="black") + 
  geom_label_repel(data=a2, aes(label=NAME, geometry=geometry), 
                   stat="sf_coordinates", size=4, nudge_x=0, nudge_y=-.5) + 
  theme_minimal() + labs(
    x="Longitude", y="Latitude",
    title="Indigenous Tribal Boundaries in AZ",
    caption="Source: Shapefile obtained using {tigris} R package, v2.0.1 
    Indigenous Tribe Shapefile obtained from AZGeo Data"
  )



```


## 5. Arizona state of patchwork
```{r}
#| label: filtering data for q5
selected_tribes <- a1 |> filter(
  startsWith(NAME, "White M") | startsWith(NAME, "San Carlos") |
    NAME == "Tohono O’odham Nation"
)

selected_counties <- azcounties |> filter(
  startsWith(NAME, "Maricopa") | startsWith(NAME, "Pinal") |  startsWith(NAME, "Pima")
)


inset_plotted_counties <- azcounties |>  filter(
  startsWith(NAME, "Maricopa") | startsWith(NAME, "Pinal") |  startsWith(NAME, "Pima") | startsWith(NAME, "Santa Cruz") | startsWith(NAME, "Yavapai") | startsWith(NAME, "Gila"))

```


```{r}
#| label: visualizing q5
inset <- ggplot(
  joined, aes(geometry=geometry, fill=total_pop_change_20_23)
) + geom_sf( aes(label=NAME, geometry=geometry), color="white")  + 
  scale_fill_gradient2(midpoint=(maxi - mini) / 2, low="dodgerblue3", high="firebrick")  +
  geom_sf(data=selected_tribes, linewidth=1, fill="lightgrey", color="black") + 
  geom_label_repel(data=selected_tribes, aes(label=NAME, geometry=geometry), inherit.aes = FALSE, nudge_x = 1, nudge_y = .5, stat="sf_coordinates", size=3) +
  coord_sf(xlim=c(-114, -110.5), ylim=c(31, 34)) + theme_minimal() + 
  labs(x="", y="") + 
  theme(
    axis.ticks.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank(),
    panel.background=element_rect(fill="darkgrey"),
    panel.grid.major=element_blank(),
    plot.margin=margin(0, 20, 0, 0),
    legend.position="None"
  )


main <- ggplot(
  joined, aes(geometry=geometry, fill=total_pop_change_20_23)
) + geom_sf( aes(label=NAME, geometry=geometry), color="white") + coord_sf(xlim=c(-122, -109), ylim=c(28, 37)) + 
  scale_fill_gradient2(midpoint=(maxi - mini) / 2, low="dodgerblue3", high="firebrick")  + 
  labs(fill = "Population Change",
       x="Longitude", y="Latitude",
       title="Resident Population Change for Counties in Arizona",
       subtitle="July 01, 2020 to July 01, 2023",
       caption="Source: Shapefile obtained using {tigris} R package, v.2.0.1\npopulation change data from the US Census Bureau\nIndigenous Tribe Shapefil obtained from AZGeo Data") +
  theme(
    legend.position=c(.2, .85),
    legend.direction="horizontal",
    legend.background = element_blank(),
    legend.box.background = element_blank()
  ) +
  guides(fill = guide_colorbar(title.position = "top"))


main + inset_element(inset,-.088, -.06, .5, .5)
``` 

